---
title: 'Reproducible Research: Peer Assessment 2'
author: "Billy Chen"
date: "05/24/2015"
output: html_document
---

## Loading and preprocessing the data

```{r}
if (!file.exists("/home/billy/workarea/coursera/reproducibleResearch/PeerAssessment2/stormData.csv.bz2")) {
  download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", 
                "/home/billy/workarea/coursera/reproducibleResearch/PeerAssessment2/stormData.csv.bz2")
}
```

```{r}
# unzip file
if (!file.exists("/home/billy/workarea/coursera/reproducibleResearch/PeerAssessment2/stormData.csv")) {
  library(R.utils)
  bunzip2("/home/billy/workarea/coursera/reproducibleResearch/PeerAssessment2/stormData.csv.bz2", "/home/billy/workarea/coursera/reproducibleResearch/PeerAssessment2/stormData.csv", remove = FALSE)
}
```

```{r}
# load data into R
storm <- read.csv("/home/billy/workarea/coursera/reproducibleResearch/PeerAssessment2/stormData.csv")
storm$year <- as.numeric(format(as.Date(storm$BGN_DATE, format = "%m/%d/%Y %H:%M:%S"), "%Y"))

mydata <- storm[c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP",  "year")]

head(mydata)

#Find unique Property Damage exponential 
unique(mydata$PROPDMGEXP)

# Sorting the property exponent data
mydata$PROPEXP[mydata$PROPDMGEXP == "K"] <- 1000
mydata$PROPEXP[mydata$PROPDMGEXP == "M"] <- 1e+06
mydata$PROPEXP[mydata$PROPDMGEXP == ""] <- 1
mydata$PROPEXP[mydata$PROPDMGEXP == "B"] <- 1e+09
mydata$PROPEXP[mydata$PROPDMGEXP == "m"] <- 1e+06
mydata$PROPEXP[mydata$PROPDMGEXP == "0"] <- 1
mydata$PROPEXP[mydata$PROPDMGEXP == "5"] <- 1e+05
mydata$PROPEXP[mydata$PROPDMGEXP == "6"] <- 1e+06
mydata$PROPEXP[mydata$PROPDMGEXP == "4"] <- 10000
mydata$PROPEXP[mydata$PROPDMGEXP == "2"] <- 100
mydata$PROPEXP[mydata$PROPDMGEXP == "3"] <- 1000
mydata$PROPEXP[mydata$PROPDMGEXP == "h"] <- 100
mydata$PROPEXP[mydata$PROPDMGEXP == "7"] <- 1e+07
mydata$PROPEXP[mydata$PROPDMGEXP == "H"] <- 100
mydata$PROPEXP[mydata$PROPDMGEXP == "1"] <- 10
mydata$PROPEXP[mydata$PROPDMGEXP == "8"] <- 1e+08
mydata$PROPEXP[mydata$PROPDMGEXP == "+"] <- 0
mydata$PROPEXP[mydata$PROPDMGEXP == "-"] <- 0
mydata$PROPEXP[mydata$PROPDMGEXP == "?"] <- 0

mydata$PROPDMG <- mydata$PROPDMG * mydata$PROPEXP

# exploring the crop exponent data
unique(mydata$CROPDMGEXP)

# Sorting the property exponent data
mydata$CROPEXP[mydata$CROPDMGEXP == "M"] <- 1e+06
mydata$CROPEXP[mydata$CROPDMGEXP == "K"] <- 1000
mydata$CROPEXP[mydata$CROPDMGEXP == "m"] <- 1e+06
mydata$CROPEXP[mydata$CROPDMGEXP == "B"] <- 1e+09
mydata$CROPEXP[mydata$CROPDMGEXP == "0"] <- 1
mydata$CROPEXP[mydata$CROPDMGEXP == "k"] <- 1000
mydata$CROPEXP[mydata$CROPDMGEXP == "2"] <- 100
mydata$CROPEXP[mydata$CROPDMGEXP == ""] <- 1
mydata$CROPEXP[mydata$CROPDMGEXP == "?"] <- 0

mydata$CROPDMG <- mydata$CROPDMG * mydata$CROPEXP

mydata$DMG <- (mydata$PROPDMG + mydata$CROPDMG) / 1000000000
finaldata <- mydata[,c("EVTYPE","FATALITIES","INJURIES","DMG", "year")]  

head(finaldata)

```

## Result

Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?


```{r}
#Summarise of Fatalities events and get the top 10 result 
FATALITIES <- aggregate(FATALITIES ~ EVTYPE, finaldata, sum)
FATALITIES <- FATALITIES[order(-FATALITIES$FATALITIES), ][1:10, ]
#Print the table
print(FATALITIES)

#Summarise of Injuries events and get the top 10 result
INJURIES <- aggregate(INJURIES ~ EVTYPE, finaldata, sum)
INJURIES <- INJURIES[order(-INJURIES$INJURIES), ][1:10, ]
#Print the table
print(INJURIES)

par(mfrow = c(1, 2))
barplot(INJURIES$INJURIES, names.arg = FATALITIES$EVTYPE, main= "Most Fatalities Events", ylab = "No. of Fatalities", las = 2, cex.names = 0.6, cex.axis = 0.6)
barplot(INJURIES$INJURIES, names.arg = INJURIES$EVTYPE, main= "Most Injuries Events", ylab = "No. of Injuries", las = 2, cex.names = 0.6, cex.axis = 0.6)

```

The most harmful with respect to population health is TORNADO

2.Across the United States, which type of events have the greatest economic consequences?
Using the same model, we get the top 10 most damaging events

```{r}
#Summarise of the top 10 most Damaging events 
DMG <- finaldata[order(-finaldata$DMG),][1:10, ]
#Print the table
print(DMG)

barplot(DMG$DMG, names.arg = DMG$EVTYPE, main= "greatest economic consequences", ylab = "Est. Cost in Billions", las = 3, cex.names = 0.6, cex.axis = 0.6)

```

The greatest economically damaging event is FLOOD

